---
title: "Use of Spatialsmooth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use of Spatialsmooth in a Example dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Installation  


```{r installation, eval=FALSE}
if(!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("Spatialsmooth")

```

### Overview  



This subsection demonstrates how to use the Spatialsmooth package.  


### Data preparation

```{r setup}
suppressMessages(library(CARD))
suppressMessages(library(Seurat))
suppressMessages(library(SeuratDisk))
suppressMessages(library(spacexr))
suppressMessages(library(Giotto))
suppressMessages(library(SPOTlight))
suppressMessages(library(Redeconve))
suppressMessages(library(AdRoit))
suppressMessages(library(SpatialDecon))
suppressMessages(library(Spatialsmooth))
```

Now let's load our example data and examine it:  

```{r loaddata}
scrna_path = 'sc_rna.h5seurat'
spatial_path = 'st_data.h5seurat'
celltype_final = 'celltype'
python_path = "your python path"   #We recommend using conda.
output_path = 'output/'     #output_path does not need to be defined in advance; save to a temporary directory.
```

The first is the deconvolution function, which has five parameters  
  
1.scrna_path Path to scRNA-seq data.  
2.spatial_path Path to spatial transcriptomics data.  
3.python_path Path to Python executable.  
4.celltype_final Column name in sc_obj@meta.data containing cell type information.  
5.output_path Directory to save output.  

```{r run_deconvolution_methods}
run_deconvolution_methods(scrna_path, spatial_path,celltype_final, python_path, output_path)
```

### smooth function
Parameters in smooth_matrices  
1.proportion_matrices A list of proportion matrices to be smoothed.  
2.pos A data frame with x and y coordinates for each pixel.  
3.n_pos_features Number of positional encoding features. Default is 16.  
4.epochs Number of training epochs. Default is 4000.  
5.batch_size Batch size for training. Default is 5.  
6.cropping A list specifying the cropping parameters. Default is list(c(0, 9), c(9, 1)).

```{r Spatialsmooth}
Sys.setenv(RETICULATE_PYTHON = "your python path")  
suppressMessages(library(Matrix))
suppressMessages(library(Seurat))
suppressMessages(library(SeuratDisk))
suppressMessages(library(keras))
suppressMessages(library(tensorflow))
suppressMessages(library(reticulate))
suppressMessages(library(Spatialsmooth))

load("output/CARD_result.Rdata")
CARD_matrix <- as.matrix(CARD_result)
load("output/seqFISH+/RCTD_result.Rdata")
RCTD_matrix <- as.matrix(RCTD_result)
load("output/SPOTlight_result.Rdata")
SPOTlight_matrix <- as.matrix(SPOTlight_result)
load("output/SpatialDWLS_result.Rdata")
SpatialDWLS_matrix <- as.matrix(SpatialDWLS_result)
load(file = 'output/pos.Rdata')

# Make the column names of the matrix names the same
colnames(SPOTlight_matrix) = gsub("\\.", "_", colnames(SPOTlight_matrix))

CARD_matrix <- CARD_matrix[,colnames(RCTD_matrix)]
SPOTlight_matrix <- SPOTlight_matrix[,colnames(RCTD_matrix)]
SpatialDWLS_matrix <- SpatialDWLS_matrix[,colnames(RCTD_matrix)]


row_names <- rownames(SpatialDWLS_matrix)
col_names <- colnames(SpatialDWLS_matrix)
SpatialDWLS_matrix <- matrix(as.numeric(SpatialDWLS_matrix), nrow = nrow(SpatialDWLS_matrix), ncol = ncol(SpatialDWLS_matrix))
rownames(SpatialDWLS_matrix) <- row_names
colnames(SpatialDWLS_matrix) <- col_names

# Prepare data
proportion_matrices <- list(CARD_matrix, RCTD_matrix, SPOTlight_matrix, SpatialDWLS_matrix)

result <- smooth_multiple_matrices(
  proportion_matrices, 
  pos,
  epochs = 4000,
  batch_size = 5,
  cropping = list(c(0, 9), c(9, 1))  # Custom Cropping Parameters
)

# Get trained models and training history
trained_model <- result$model
training_history <- result$history
# smoothed_matrices <- result$smoothed_matrices
smooth_result <- result$smoothed_matrices
plot(training_history)


smooth_result
save(smooth_result, file = "output/smooth_result.RData")
```

### Add new method function
```{r Add method}
run_menthod_name <- function(scrna_path, spatial_path, celltype_final, output_path) {
    #Deconvolution logic
    save(deconvolution_result, file = file.path(output_path, "deconvolution_result.Rdata"))
}
```

```{r Define registration information}
registered_methods <- list(
  Deconvolution_method = list(
    function_name = "run_method",  # Function name (string)
    use_python = FALSE,              # Whether to use Python
    packages = c("reticulate")      # Additional R packages required,such as, reticulate
  )
)
```
